---
title: "lab5_lorenestasiuk"
author: "Lorene Stasiuk"
date: "9/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("tidyverse")
```

```{r, 6_13_data}
Confirmed_State_6_13<-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/06-13-2020.csv")) %>%
  filter (Country_Region == "US") %>% 
  group_by(Province_State, Country_Region) %>% 
  summarise(Confirmed = sum(Confirmed)) 
```

```{r, check_6_13_data}
str(Confirmed_State_6_13)
```

```{r, 9_13_data}
Confirmed_State_9_13 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-13-2020.csv")) %>% 
  filter (Country_Region == "US") %>% 
  group_by(Province_State, Country_Region) %>% 
  summarise(Confirmed = sum(Confirmed)) 
```

```{r, check_9_13_data}
str(Confirmed_State_9_13)

Confirmed_State_9_13 <- Confirmed_State_9_13 %>% 
  filter(Province_State != "Recovered") 
```

### Exercise 1
```{r, exercise1}
### Please see datavisualization_lorenestasiuk.html
```


### Exercise 2
```{r, exercise2, fig.width = 5, fig.height = 10}
### Revise the above example for joining tables using 6/13/2020 and 9/13/2020. Plot the data as a bar plot. Make sure your graph looks nice in the report by adjusting the height in the R code chunk header.

Confirmed_State_6_13_9_13_joined <- full_join(Confirmed_State_6_13,
      Confirmed_State_9_13, by = c("Province_State"))
head(Confirmed_State_6_13_9_13_joined)

Confirmed_State_6_13_9_13_joined <- full_join(Confirmed_State_6_13,
      Confirmed_State_9_13, by = c("Province_State")) %>% 
      rename(Confirmed_6_13_2020 = "Confirmed.x", Confirmed_9_13_2020 = "Confirmed.y") %>% 
      select(-Country_Region.x, -Country_Region.y) %>% 
      replace_na(list(Confirmed_6_13_2020 = 0))
head(Confirmed_State_6_13_9_13_joined)

# which(is.na(Confirmed_State_6_13_9_13_joined))

Confirmed_State_6_13_9_13_joined_long <- Confirmed_State_6_13_9_13_joined %>% 
              pivot_longer(-c(Province_State),
                            names_to = "Date", values_to = "Confirmed")

ggplot(Confirmed_State_6_13_9_13_joined_long, aes(x = Confirmed,  y = Province_State))  + geom_bar(stat="identity",aes(color = Date))
```

### Exercise 3
```{r, exercise3, fig.width = 5, fig.height = 10}
### Add more informative x and y labels and a title to the above graph in Ex2.

ggplot(Confirmed_State_6_13_9_13_joined_long, aes(x = Confirmed,  y = Province_State))  + geom_bar(stat="identity",aes(color = Date)) +  labs(title="COVID-19 Confirmed Cases in US",
        x ="Number of Confirmed Cases", y = "State/Province in US")
```

```{r}
library(lubridate)
```

```{r, exercise, fig.width = 5, fig.height = 10}
### Use Mutate to make a new column with deaths/confirmed cases (Hint: this might be best done after joining the tables, but before pivoting longer).

time_series_confirmed <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")

head(time_series_confirmed)

time_series_confirmed_long <- time_series_confirmed %>% 
               pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                            names_to = "Date", values_to = "Confirmed") 

head(time_series_confirmed_long)

time_series_deaths <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")%>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")

time_series_deaths_long <- time_series_deaths %>% 
               pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                            names_to = "Date", values_to = "Deaths") 
head(time_series_deaths_long)

time_series_deaths_long <- time_series_deaths_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".") %>% 
  select(Key, Deaths)

time_series_confirmed_long <- time_series_confirmed_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".", remove = FALSE)
head(time_series_confirmed_long)

time_series_long_joined <- full_join(time_series_confirmed_long,
              time_series_deaths_long, by = c("Key")) %>% 
              select(-Key)
head(time_series_long_joined)

#which(is.na(time_series_long_joined$Confirmed))

#which(is.na(time_series_long_joined$Deaths))

time_series_long_joined$Date <- mdy(time_series_long_joined$Date)
```

### Exercise 4
```{r, exercise4, fig.width = 5, fig.height = 10}
### Using the time series data. Make a plot of the total number of confirmed deaths per day worldwide
time_series_long_joined %>% 
  group_by(Date) %>% 
  summarise_at(c("Confirmed", "Deaths"), sum) %>%
    ggplot(aes(x = Date,  y = Deaths)) + 
    geom_point() +
    geom_line() +
    ggtitle("Worldwide COVID-19 Deaths")
```

### Exercise 5
```{r, exefcise 5, fig.width = 5, fig.height = 10}
### Use Mutate to make a new column with deaths/confirmed cases (Hint: this might be best done after joining the tables, but before pivoting longer).
time_series_long_joined %>% 
  mutate(ratio = Deaths/Confirmed)

time_series_long_joined_counts <- time_series_long_joined %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long, Date, Deaths/Confirmed),
               names_to = "Report_Type", values_to = "Counts")
str(time_series_long_joined_counts)
```

### Exercise 6
```{r, exercise6, fig.width = 5, fig.height = 10}
### Plot US deaths/confirmed cases per day.

time_series_long_joined %>% 
  group_by(Country_Region,Date) %>% 
  summarise_at(c("Confirmed", "Deaths"), sum) %>% 
  filter (Country_Region == "US") %>% 
    ggplot(aes(x = Date,  y = Deaths/Confirmed)) + 
    geom_point() +
    geom_line() +
    ggtitle("US COVID-19 Deaths")
```

### Exercise 7
```{r, exercise7, fig.width = 5, fig.height = 10}
### Make a single graph with from the times series data with the 10 countries with the hightest death total.
Confirmed <- time_series_long_joined %>% arrange(desc(Confirmed)) %>% slice(1:10)

time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Deaths"), sum) %>% 
    filter (Country_Region %in% c("Brazil","UK","Italy", 
                                "Mexico", "US","Spain","France","Russia","Germany","China")) %>% 
    ggplot(aes(x = Date,  y = Deaths, color = Country_Region)) + 
    geom_point() +
    geom_line() +
    ggtitle("COVID-19 Deaths")
```

### Exercise 8
```{r, exercise8, fig.width = 5, fig.height = 10}
### Using Facet_Wraps make a graph from the times series data with the 10 countries with the hightest death total. Format the table to look nice in the report by changing the number of rows in your graph and/or adjusting the height in the R code chunk header.

time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Deaths"), sum) %>% 
    filter (Country_Region %in% c("Brazil","UK","Italy", 
                                "Mexico", "US","Spain","France","Russia","Germany","China")) %>% 
    ggplot(aes(x = Date,  y = Deaths, color = Country_Region)) + 
    geom_point() +
    geom_line() +
  facet_wrap(~Country_Region, ncol=2, scales="free_y") +
    ggtitle("COVID-19 Deaths")
```

### Exercise 9
#### I tried to do this question, but I ran into a number of problems. My code is shown below. 
```{r, load_us_confirmed, fig.width = 5, fig.height = 10}
### Using Facet_Wraps make a graph from the times series data for confirmed cases for all US states and territories. Format the table to look nice in the report by changing the number of rows in your graph and/or adjusting the height in the R code chunk header.

us_time_series_confirmed <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv"))

head(us_time_series_confirmed)
```

```{r}
us_time_series_deaths <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv"))

head(us_time_series_deaths)
```
#### This is the chunk gave the error: Error: Can't subset columns that don't exist. x Column `Population` doesn't exist. I have commented it out to show what I had done. 
```{r}
#us_time_series_confirmed_long <- us_time_series_confirmed %>% 
#               pivot_longer(-c(UID, iso2, iso3, code3, FIPS, Admin2, #Province_State, Country_Region, Lat, Long_, Combined_Key, Population),
#                            names_to = "Date", values_to = "Confirmed") 

#head(us_time_series_confirmed_long)

```

```{r}
#us_time_series_deaths_long <- us_time_series_deaths %>% 
#               pivot_longer(-c(UID, iso2, iso3, code3, FIPS, Admin2, Combined_Key, Province_State, Country_Region, Lat, Long_),
#                            names_to = "Date", values_to = "Confirmed") 
#head(us_time_series_deaths_long)
```

### Exercise 10
```{r, exercise10, fig.width = 5, fig.height = 10}
### Apply a theme other than the one used in the Data Carpentries tutorial to the graph in Ex8.

theme_set(theme_bw())

time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Deaths"), sum) %>% 
    filter (Country_Region %in% c("Brazil","UK","Italy", 
                                "Mexico", "US","Spain","France","Russia","Germany","China")) %>% 
    ggplot(aes(x = Date,  y = Deaths, color = Country_Region)) +
    geom_point() +
    geom_line() +
  facet_wrap(~Country_Region, ncol=2, scales="free_y") +
    ggtitle("COVID-19 Deaths") + theme(
  panel.background = element_rect(fill = "lightblue",
                                colour = "lightblue",
                                size = 0.5, linetype = "solid"),
  panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "white"), 
  panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                colour = "white")
  )
```

Application written in R [@RCoreTeam] using the Shiny framework [@Chang2015].

# REFERENCES
```
@Misc{Chang2015,
  Title                    = {shiny: Web Application Framework for R. R package version 0.12.1},

  Author                   = {Chang, W. and Cheng, J. and Allaire, JJ. and Xie, Y. and McPherson, J. },
  Year                     = {2015},

  Type                     = {Computer Program},
  Url                      = {http://CRAN.R-project.org/package=shiny}
}

@Article{RCoreTeam,
  Title                    = {R: A Language and Environment for Statistical Computing},
  Author                   = {{R Core Team}},
  Year                     = {2015},

  Type                     = {Journal Article},
  Url                      = {http://www.R-project.org}
}
```
