---
title: "lab6report_lorenestasiuk"
author: "Lorene Stasiuk"
date: "9/28/2020"
output: html_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, load_packages, echo = FALSE, message = FALSE, warning = FALSE}
library("tidyverse")
library(maps)
library(mapdata)
library(lubridate)
library(viridis)
library(wesanderson)
library(plotly)
library(RColorBrewer)
library(gganimate)
library(transformr)
# To display only colorblind-friendly brewer palettes, specify the option colorblindFriendly = TRUE as follow:
# display.brewer.all(colorblindFriendly = TRUE)
```

# Comparing the world confirmed cases from six months ago and now

## World confirmed cases from six months ago (3/26/20)
```{r, daily_report1_world, echo = FALSE, message = FALSE, warning = FALSE}
# loading data

daily_report1 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-26-2020.csv")) %>% 
  rename(Long = "Long_") 

# graphing data

ggplot(daily_report1, aes(x = Long, y = Lat, size = Confirmed/1000)) +
    borders("world", colour = NA, fill = "grey90") +
    theme_bw() +
    geom_point(shape = 21, color='blue', fill='red', alpha = 0.5) +
    labs(title = 'World COVID-19 Confirmed Cases - 3/26/20',x = '', y = '',
        size="Cases (x1000))") +
    theme(legend.position = "right") +
    coord_fixed(ratio=1.5)
```

## World confirmed cases now (9/26/20)
```{r, daily_report2_world, echo = FALSE, message = FALSE, warning = FALSE}
### load data from 9.26.20
daily_report_2 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  rename(Long = "Long_") 
    
### we adjust the confirmed cases by dividing by 1 million

ggplot(daily_report_2, aes(x = Long, y = Lat, size = Confirmed/1000000)) +
    borders("world", colour = NA, fill = "grey90") +
    theme_bw() +
    geom_point(shape = 21, color='blue', fill='red', alpha = 0.5) +
    labs(title = 'World COVID-19 Confirmed Cases - 9/26/20',x = '', y = '',
        size="Cases (x1000000))") +
    theme(legend.position = "right") +
    coord_fixed(ratio=1.5)
```
## Visualizing the number of confirmed cases in the US by state from six months ago (3/26/20) and now (9/26/20)

```{r, daily_report_confirmed_us, echo = FALSE, message = FALSE, warning = FALSE}
Confirmed_State_3_26<-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-26-2020.csv")) %>%
  filter (Country_Region == "US") %>% 
  group_by(Province_State, Country_Region) %>% 
  summarise(Confirmed = sum(Confirmed)) 

Confirmed_State_9_26 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  filter (Country_Region == "US") %>% 
  group_by(Province_State, Country_Region) %>% 
  summarise(Confirmed = sum(Confirmed)) 

Confirmed_State_9_26 <- Confirmed_State_9_26 %>% 
  filter(Province_State != "Recovered") 

Confirmed_State_3_26_9_26_joined <- full_join(
      Confirmed_State_9_26,Confirmed_State_3_26, by = c("Province_State"))

Confirmed_State_3_26_9_26_joined <- full_join(
      Confirmed_State_9_26, Confirmed_State_3_26, by = c("Province_State")) %>% 
      rename(Confirmed_3_26_2020 = "Confirmed.x", Confirmed_9_26_2020 = "Confirmed.y") %>% 
      select(-Country_Region.x, -Country_Region.y) %>% 
      replace_na(list(Confirmed_3_26_2020 = 0))

# which(is.na(Confirmed_State_6_13_9_13_joined))

Confirmed_State_3_26_9_26_joined_long <- Confirmed_State_3_26_9_26_joined %>% 
              pivot_longer(-c(Province_State),
                            names_to = "Date", values_to = "Confirmed")

ggplotly(
  ggplot(Confirmed_State_3_26_9_26_joined_long, aes(x = Confirmed,  y = Province_State))  + geom_bar(stat="identity",aes(color = Date)) +  labs(title="COVID-19 Confirmed Cases in US",
        x ="Number of Confirmed Cases", y = "State/Province in US")
 )
```
# Comparing the US confirmed cases from six months ago and now

## Looking at the US six months ago
```{r, daily_report1_us, echo = FALSE, message = FALSE, warning = FALSE}
daily_report_3 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Country_Region == "US") %>% 
  filter (!Province_State %in% c("Alaska","Hawaii", "American Samoa",
                  "Puerto Rico","Northern Mariana Islands", 
                  "Virgin Islands", "Recovered", "Guam", "Grand Princess",
                  "District of Columbia", "Diamond Princess")) %>% 
  filter(Lat > 0)
ggplot(daily_report_3, aes(x = Long, y = Lat, size = Confirmed/1000)) +
    borders("state", colour = "black", fill = "grey90") +
    theme_bw() +
    geom_point(shape = 21, color='blue', fill='purple', alpha = 0.5) +
    labs(title = 'COVID-19 Confirmed Cases in the US - 3/26/20', x = '', y = '',
        size="Cases (x1000))") +
    theme(legend.position = "right") +
    coord_fixed(ratio=1.5)
```

## US confirmed cases now (9/26/20)
```{r, daily_report2_us, echo = FALSE, message = FALSE, warning = FALSE}
daily_report_3 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Country_Region == "US") %>% 
  filter (!Province_State %in% c("Alaska","Hawaii", "American Samoa",
                  "Puerto Rico","Northern Mariana Islands", 
                  "Virgin Islands", "Recovered", "Guam", "Grand Princess",
                  "District of Columbia", "Diamond Princess")) %>% 
  filter(Lat > 0)
ggplot(daily_report_3, aes(x = Long, y = Lat, size = Confirmed/1000)) +
    borders("state", colour = "black", fill = "grey90") +
    theme_bw() +
    geom_point(shape = 21, color='blue', fill='purple', alpha = 0.5) +
    labs(title = 'COVID-19 Confirmed Cases in the US - 3/26/20', x = '', y = '',
        size="Cases (x1000))") +
    theme(legend.position = "right") +
    coord_fixed(ratio=1.5)
```

# Comparing the US confirmed cases from six months ago and now based on county

## Looking at the US six months ago based on county (3/26/20)
```{r, county_report1, echo = FALSE, message = FALSE, warning = FALSE}
# Get and format the covid report data
report_03_26_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  unite(Key, Admin2, Province_State, sep = ".") %>% 
  group_by(Key) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Key = tolower(Key))

# get and format the map data
us <- map_data("state")
counties <- map_data("county") %>% 
  unite(Key, subregion, region, sep = ".", remove = FALSE)
# Join the 2 tibbles
state_join <- left_join(counties, report_03_26_2020, by = c("Key"))
# sum(is.na(state_join$Confirmed))
ggplot(data = us, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  # Add data layer
  borders("state", colour = "black") +
  geom_polygon(data = state_join, aes(fill = Confirmed)) +
  scale_fill_gradientn(colors = brewer.pal(n = 5, name = "PuRd"),
                       breaks = c(1, 10, 100, 1000, 10000, 100000),
                       trans = "log10", na.value = "White") +
  ggtitle("Number of Confirmed Cases by US County") +
  theme_bw() 
```

## Looking at the US now based on county (9/26/20)
```{r, county_report2, echo = FALSE, message = FALSE, warning = FALSE}
# Get and format the covid report data
report_09_26_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  unite(Key, Admin2, Province_State, sep = ".") %>% 
  group_by(Key) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Key = tolower(Key))

# get and format the map data
us <- map_data("state")
counties <- map_data("county") %>% 
  unite(Key, subregion, region, sep = ".", remove = FALSE)
# Join the 2 tibbles
state_join <- left_join(counties, report_09_26_2020, by = c("Key"))
# sum(is.na(state_join$Confirmed))
ggplot(data = us, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  # Add data layer
  borders("state", colour = "black") +
  geom_polygon(data = state_join, aes(fill = Confirmed)) +
  scale_fill_gradientn(colors = brewer.pal(n = 5, name = "PuRd"),
                       breaks = c(1, 10, 100, 1000, 10000, 100000),
                       trans = "log10", na.value = "White") +
  ggtitle("Number of Confirmed Cases by US County") +
  theme_bw() 
```

# Comparing the Florida confirmed cases from six months ago and now using interactive graphs

## Confirmed cases in Florida from six months ago (3/26/20)
```{r, fl_data1, echo = FALSE, message = FALSE, warning = FALSE}
daily_report_ex4 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Province_State == "Florida") %>% 
  group_by(Admin2) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Admin2 = tolower(Admin2))

us_ex4 <- map_data("state")

fl_us <- subset(us_ex4, region == "florida")

counties <- map_data("county")

fl_county <- subset(counties, region == "florida")

state_join_ex4 <- left_join(fl_county, daily_report_ex4, by = c("subregion" = "Admin2")) 

# plot state map

ggplotly(
  ggplot(data = fl_county, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = state_join_ex4, aes(fill = Confirmed), color = "black") +
    scale_fill_gradientn(colours = 
                         wes_palette("FantasticFox1", 100, type = "continuous")) +
  ggtitle("COVID-19 Cases in FL - 9/26/20") +
# Cleaning up the graph
  labs(x=NULL, y=NULL) +
  theme(panel.border = element_blank()) +
  theme(panel.background = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(axis.text = element_blank())
)
```

## Confirmed cases in Florida now (9/26/20)
```{r, fl_data2, echo = FALSE, message = FALSE, warning = FALSE}
### Make an interactive plot using a state of your choosing using a theme different from used in the above examples.

### Use Florida data

daily_report_ex4 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Province_State == "Florida") %>% 
  group_by(Admin2) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Admin2 = tolower(Admin2))

us_ex4 <- map_data("state")

fl_us <- subset(us_ex4, region == "florida")

counties <- map_data("county")

fl_county <- subset(counties, region == "florida")

state_join_ex4 <- left_join(fl_county, daily_report_ex4, by = c("subregion" = "Admin2")) 

# plot state map

ggplotly(
  ggplot(data = fl_county, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = state_join_ex4, aes(fill = Confirmed), color = "black") +
    scale_fill_gradientn(colours = 
                         wes_palette("FantasticFox1", 100, type = "continuous")) +
  ggtitle("COVID-19 Cases in FL - 9/26/20") +
# Cleaning up the graph
  labs(x=NULL, y=NULL) +
  theme(panel.border = element_blank()) +
  theme(panel.background = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(axis.text = element_blank())
)
```
Application written in R [@RCoreTeam] using the Shiny framework [@Chang2015]. Data aquired from [@JohnsHopkinsUniversity].

# REFERENCES

